The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-13 paper/paper.Rmd
Change data trimming for eda.

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, 
               message = FALSE,
               cache = TRUE, 
               cache.path = "paper_cache/",
               fig.path = "figures/",
               comment = "#>",
               fig.align = 'center', #fig.show = 'hold',
               out.width = "70%")

@@

The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent. 
Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor. Since we do modeling, it is important to look at the number of observations from each respondent. Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies. Moreover, each number of observations seems to have an impact on the slope. Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations. Accordingly, we trim the data to only include $5^{th} - 12^{th}$ percentile.

@@

```{r gap-gender, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males. Females tend to earn fewer than males. For example, in 2018 females earn about 78 cents for every \\$1 earned by males. Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class."}
```

>>>>

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, 
               message = FALSE,
               cache = TRUE, 
               cache.path = "paper_cache/",
               fig.path = "figures/",
               comment = "#>",
               fig.align = 'center', 
               out.width = "70%")

@@

The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent. 
Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor. Since we do modeling, it is important to look at the number of observations from each respondent. Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies. Moreover, each number of observations seems to have an impact on the slope. Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations. Accordingly, we trim the data to only include $5^{th} - 95^{th}$ percentile.

@@

```{r gap-gender, fig.cap="A) Number of observations by gender and by year. There are more males participated in survey than females. The number of observations decreased over the years. B) Gender wage gap or the median wages of females relatives to median wages of males. Females tend to earn fewer than males. For example, in 2018 females earn about 78 cents for every \\$1 earned by males. Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.", out.width="90%"}
```

### 2021-09-13 paper/paper.tex
Change data trimming for eda.

The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent.
Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor. Since we do modeling, it is important to look at the number of observations from each respondent. Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies. Moreover, each number of observations seems to have an impact on the slope. Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations. Accordingly, we trim the data to only include \(5^{th} - 12^{th}\) percentile.

@@

{\centering \includegraphics[width=0.7\linewidth]{figures/gap-gender-1}

@@

\caption{Gender wage gap or the median wages of females relatives to median wages of males. Females tend to earn fewer than males. For example, in 2018 females earn about 78 cents for every \$1 earned by males. Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.}\label{fig:gap-gender}
\end{figure}

>>>>

The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent.
Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor. Since we do modeling, it is important to look at the number of observations from each respondent. Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies. Moreover, each number of observations seems to have an impact on the slope. Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations. Accordingly, we trim the data to only include \(5^{th} - 95^{th}\) percentile.

@@

{\centering \includegraphics[width=0.9\linewidth]{figures/gap-gender-1}

@@

\caption{A) Number of observations by gender and by year. There are more males participated in survey than females. The number of observations decreased over the years. B) Gender wage gap or the median wages of females relatives to median wages of males. Females tend to earn fewer than males. For example, in 2018 females earn about 78 cents for every \$1 earned by males. Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.}\label{fig:gap-gender}
\end{figure}
